### 2015-12-14 paper/4-performance.tex
Proofread section 4.4

\subsection{Memory Usage}
Apart from simulation execution time, memory usage during simulation is also of great importance.
While execution time only becomes a problem if it takes way too long, coming short only a bit of memory can make simulation infeasible.
We therefore also investigate memory usage of different synchronization protocols.

We do not tackle the problem of states that become too large for a single machine to hold.
This problem can be mitigated by distribution over multiple machines, which neither dxex or adevs support.
Comparison of memory usage should thus only be seen in the context of which models are feasible to be simulated by the tool.

\subsubsection{Remarks}
Both dxex and adevs use \texttt{tcmalloc} as memory allocator.
Additionally, dxex uses memory pools to further reduce the frequency of expensive system calls (\textit{e.g.}, malloc and free).
\texttt{tcmalloc} only gradually releases memory back to the OS, whereas our pools will not do so at all.
If memory has been allocated once, it is, at least from a performance point of view, better to keep that memory in the pool indefinitely.
Due to our motivation for memory usage analysis, we will only measure peak allocation.
Profiling is done using Valgrind's massif tool~\cite{Nethercote:2007:VFH:1273442.1250746}.

\subsubsection{Results}
Figure~\ref{fig:memory} shows the memory used by each different benchmark.
Results are in megabytes, and show the total memory footprint of the running application (\textit{i.e.}, text, stack, and heap).

Unsurprisingly, optimistic synchronization results show very high memory usage due to the saved states.
Note the logarithmic scale that was used for this reason.
Also, results for optimistic synchronization vary heavily depending on thread scheduling by the operating system, as this influences the drift between different nodes.
Comparing similar approaches though, we notice that dxex and adevs have very similar memory use.

>>>>

\subsection{Memory Usage}
Apart from simulation execution time, memory usage during simulation is also of great importance.
While execution time only becomes a problem if it takes way too long, coming short only a bit of memory can make simulation infeasible.
We therefore also investigate memory usage of different synchronization protocols, and again compare to adevs.

We do not tackle the problem of states that become too large for a single machine to hold.
This problem can be mitigated by distribution over multiple machines, which neither dxex or adevs support.

\subsubsection{Remarks}
Both dxex and adevs use \texttt{tcmalloc} as memory allocator, allowing for thread-local allocation.
Additionally, dxex uses memory pools to further reduce the frequency of expensive system calls (\textit{e.g.}, malloc and free).
\texttt{tcmalloc} only gradually releases memory back to the OS, whereas our pools will not do so at all.
Due to our motivation for memory usage analysis, we will only measure peak allocation.
Profiling is done using Valgrind's massif tool~\cite{Nethercote:2007:VFH:1273442.1250746}.

\subsubsection{Results}
Figure~\ref{fig:memory} shows the memory used by the different benchmarks.
Results are in megabytes, and show the total memory footprint of the running application (\textit{i.e.}, text, stack, and heap).

Unsurprisingly, optimistic synchronization results show very high memory usage due to the saved states.
Note the logarithmic scale that was used for this reason.
Optimistic synchronization results vary heavily depending on thread scheduling by the operating system, as this influences the drift between nodes.
Comparing similar approaches, we notice that dxex and adevs have very similar memory use.
